---
title: "Causal Impact of Adopting the Three Points per Victory based System on the Competitive Balance of Spanish Football League"
author: "Cesar Soto Valero"
fig_caption: TRUE
date: "Mar 5, 2018"
output:
  html_document:
    theme: united
    highlight: tango
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 3
    number_sections: true
    fig_retina: 2
  html_notebook: default
  pdf_document: default
---


```{r echo=FALSE, message=FALSE, warning=FALSE}
# import libraries
library(tidyverse)
library(dtw)
library(dtwclust)
library(ggpubr)
library(ggridges)
library(ggcorrplot)
library(caret)
library(factoextra)
library(Boruta)
library(randomForest)
library(DT)
library(ggridges)
library(RANN)
library(car)
library(MASS)
library(gplots)
library(multcomp)
library(psych)
library(plotly)
library(lubridate)
library(dendextend)
library(broom)
library(NbClust)
library(fpc)
library(devtools)
library(CausalImpact)

theme_set(theme_bw())


load(file = "data/data.RData")


```

# Introduction

The competition format of the [Spanish Football League](https://en.wikipedia.org/wiki/La_Liga) follows the usual double round-robin format. During the course of a season, which lasts from August to May, each club plays every other club twice, once at home and once away, for a total of 38 matchdays. Teams receive three points for a win, one point for a draw, and no points for a loss. Teams are ranked by total points, with the highest-ranked club at the end of the season crowned champion.

# Standard Measures of Teams' Performance

## WR

The Win Ratio ($\text{WR}$) of a team $t$ as follow

$$
\begin{equation}
\text{WR}_t = \dfrac{W_t}{W_t + L_t + D_t} 
\end{equation}
$$


where $W = \text{Games Win}$, $L = \text{Games Lost}$ and $D = \text{Ties}$


# Standard Measures of CB

## SD

Is the ratio of the actual standard deviation to an idealized standard deviation. The ideal ratio is 1. The higher the ratio, the more the actual spread diverges from the ideal one and hence the worse the competitive balance.

$$
\begin{equation}
\text{SD} = \sqrt{\sum_{i=1}^{n}\dfrac{w_i - \text{avg_wins}}{n}} 
\end{equation}
$$

where $W_i$ is the proportion of wins of team $i$.


## RE

Relative entropy

$$
\begin{equation}
\text{AC} = \dfrac{H}{H_M}
\end{equation}
$$

with


$$
\begin{equation}
\text{H} = - \sum_{i=1}^{n}p_i \lg_2p_i 
\end{equation}
$$

where

$i$ = team; $n$ = total nr of teams; $p_i$ = the proportion of the league victories of team $i$.
$H_M$ = maximal entropy: every team has the same share of victories: $p_i = 1/n$


## NAMSI

When all teams win half of their games or all games are ended in a tie complete balance occurs. The standard deviation of the season equals zero since the win percentage of all teams is 0,5. Hence the NAMSI will equal zero. If there is absolute certainty about who wins the game the seasonal standard deviation equals the certainty standard deviation and so NAMSI equals 1. This is the situation of complete imbalance. So the NAMSI ranges between 0 and 1. When comparing two seasons or two countries a higher NAMSI indicates a higher seasonal imbalance.

$$
\begin{equation}
\text{NAMSI} = \sqrt{\dfrac{\sum_{i=1}^{n}(w_i - 1/2)^2}{\sum_{i=1}^{n}(w_{i,\text{max}} - 1/2)^2}}
\end{equation}
$$

## HICB


$$
\begin{equation}
\text{HICB} = 100n\sqrt{s_{i}^2}
\end{equation}
$$
where $s_i$ is the ratio of points scored in a season by team $i$. 



## HHIN

Is a normalised version of HHI (HHIN) 

$$
\begin{equation}
\text{HHIN} = \dfrac{HHI - 1/n}{1-1/n}
\end{equation}
$$


## AC

Adjusted Churn is calculated as:


$$
\begin{equation}
\text{C}_t = \dfrac{1}{n}\sqrt{|c_{i,t} - c_{i,t-1}|}
\end{equation}
$$

where $s_i$ is the ranking of team $i$ in season $t$. Then the adjusted Churn is defined as

$$
\begin{equation}
\text{AC} = \dfrac{C_t}{C_{t,max}}
\end{equation}
$$

where $C_{t,max}$ is the maximum value of $C_t$ given a league of n teams.





# Analysis of CB

## SD analysis

In general, the highesr is SD the lower is the CB (and the higher is the *competitive unequality*)

```{r echo=FALSE, message=FALSE, warning=FALSE}

df <-
  data %>% group_by(Season, Team) %>% 
  summarise(
    Ws = sum(Game_Result == "G"),
    Ls = sum(Game_Result == "P"),
    Ts = sum(Game_Result == "E")
    ) %>% 
  mutate(Win_Ratio = Ws / (Ws + Ls + Ts))

# number of teams
n_teams <- data %>%
  group_by(Season, Team) %>%
  summarise(Total_Points = sum(Points)) %>%
  ungroup() %>%
  group_by(Season) %>%
  summarise(Count = n())

tmp <- full_join(df, n_teams, by = "Season")

SD <- tmp %>%
  ungroup() %>%
  group_by(Season, Team) %>%
  mutate(SD = (Win_Ratio - 0.5) ^ 2) %>%
  ungroup() %>%
  group_by(Season) %>%
  summarise(SD = sqrt(mean(SD)), Teams_Count = sum(Count/Count))


SD %>% ggplot(aes(Season, SD)) +
  geom_line()+
  geom_smooth(span = 0.2)

```

## RE analysis

The maximun entropy for 20 teams is: 4.32192809488736


```{r echo=FALSE, message=FALSE, warning=FALSE}

fun_HM <- function(n) {
  i <-  1
  HM <- 0
  for (i in 1:n) {
    HM <-  HM + (-1 * sum((1 / n) * log2((1 / n))))
  }
  return(dplyr::data_frame(HM = HM))
}

# cat(paste("The maximun entropy for 20 teams is", fun_HM(20)))

df <-
  data %>% group_by(Season, Team) %>% 
  summarise(
    Ws = sum(Game_Result == "G"),
    Ls = sum(Game_Result == "P"),
    Ts = sum(Game_Result == "E")
    ) %>% 
  mutate(Win_Ratio = Ws / (Ws + Ls + Ts)) %>% 
  ungroup() %>% 
  group_by(Season) %>% 
  summarise(H = -1*sum(Win_Ratio*log2(Win_Ratio)))


# number of teams
n_teams <- data %>%
  group_by(Season, Team) %>%
  summarise(Total_Points = sum(Points)) %>%
  ungroup() %>%
  group_by(Season) %>%
  summarise(Count = n()) 


hm <- n_teams %>% ungroup() %>% group_by(Season) %>% do(fun_HM(n =.$Count))

RE <- full_join(df,hm, by = "Season") %>% mutate(RE = H/HM)

RE %>% ggplot(aes(Season, RE)) +
  geom_line()+
  geom_smooth(span = 0.2)

```



## NAMSI analysis

```{r echo=FALSE, message=FALSE, warning=FALSE}

# auxiliar function that returns the denominator
my_fun <- function(n) {
  n <- n-1
  i <- 1
  result <- 0
  while (i < n-1) {
    result <- result + ((n - i) / (n - 1) - 0.5) ^ 2
    i <- i +1
  }
  return(dplyr::data_frame(result = result))
}


# number of teams
n_teams <- data %>%
  group_by(Season, Team) %>%
  summarise(Total_Points = sum(Points)) %>%
  ungroup() %>%
  group_by(Season) %>%
  summarise(Count = n()) 

# denominator
denominator <- n_teams %>% 
  ungroup() %>%
  group_by(Season) %>%
  do(my_fun(n=.$Count)) %>% 
  mutate(denominator = sqrt(result))

# numerator
numerator <- tmp %>%
  ungroup() %>%
  group_by(Season) %>%
  summarise(numerator = sqrt(sum((Win_Ratio - 0.5) ^ 2)))


NAMSI <- full_join(denominator, numerator, by = "Season") %>% 
  mutate(NAMSI = numerator/denominator)

NAMSI %>% ggplot(aes(Season, NAMSI)) +
  geom_line() +
  geom_smooth(span = 0.2)

```

## HICB analysis

```{r echo=FALSE, message=FALSE, warning=FALSE}

# S squared
df <-
  data %>% group_by(Season, Team) %>%
  summarise(Total_Points = sum(Points)) %>%
  mutate(S_quared = Total_Points ^ 2)


# number of teams
n_teams <- data %>%
  group_by(Season, Team) %>%
  summarise(Total_Points = sum(Points)) %>%
  ungroup() %>%
  group_by(Season) %>%
  summarise(Count = n())

# total points
total_points <- data %>%
  group_by(Season, Team) %>%
  summarise(Total_Points = sum(Points)) %>%
  ungroup() %>%
  group_by(Season) %>%
  summarise(Total_Points_Season = sum(Total_Points))

tmp <- full_join(df, n_teams, by = "Season")
tmp <- full_join(tmp, total_points, by = "Season")

HICB <- tmp %>%
  ungroup() %>%
  group_by(Season) %>%
  summarise(HICB = sum(Count / Count) * 100 * (sum((
    Total_Points / Total_Points_Season
  ) ^ 2)))

HICB %>% ggplot(aes(Season, HICB)) +
  geom_line()+
  geom_smooth(span = 0.2)

```


## HHIN analysis

```{r echo=FALSE, message=FALSE, warning=FALSE}

# S squared
df <-
  data %>% group_by(Season, Team) %>%
  summarise(Total_Points = sum(Points)) %>%
  mutate(S_quared = Total_Points ^ 2)


# number of teams
n_teams <- data %>%
  group_by(Season, Team) %>%
  summarise(Total_Points = sum(Points)) %>%
  ungroup() %>%
  group_by(Season) %>%
  summarise(Count = n())

# total points
total_points <- data %>%
  group_by(Season, Team) %>%
  summarise(Total_Points = sum(Points)) %>%
  ungroup() %>%
  group_by(Season) %>%
  summarise(Total_Points_Season = sum(Total_Points))

tmp <- full_join(df, n_teams, by = "Season")
tmp <- full_join(tmp, total_points, by = "Season")


HHIN <- tmp %>%
  ungroup() %>%
  group_by(Season) %>%
  summarise( HHIN = 100*(((sum((Total_Points / Total_Points_Season) ^ 2)) - 1/(sum(Count / Count))) / (1 - 1/(sum(Count / Count)))))


HHIN %>% ggplot(aes(Season, HHIN)) +
  geom_line()+
  geom_smooth(span = 0.2)




```


## AC analysis

```{r echo=FALSE, message=FALSE, warning=FALSE}
# S squared
df <-
  data %>% group_by(Season, Team) %>%
  summarise(Total_Points = sum(Points)) %>% 
  filter(Team == "Barcelona")

i <- 2
c <- c(1:nrow(df))
for (i in 1:nrow(df)) {
  c[i-1] <- abs(df$Total_Points[i] - df$Total_Points[i - 1])
}

plot(c, type = "l", xlab = "Seasons", ylab = "AC", main = "Points difference of Barcelona across seasons" )

```


# Clustering of Teams based on Teams' Performance

```{r echo=FALSE, message=FALSE, warning=FALSE}
# data preprocessing
df <-
  data %>% 
  group_by(Season, Team) %>% 
  summarise(
    Ws = sum(Game_Result == "G"),
    Ls = sum(Game_Result == "P"),
    Ts = sum(Game_Result == "E")
    ) %>% 
  mutate(Win_Ratio = Ws / (Ws + Ls + Ts)) %>% 
  dplyr::select(Season, Team, Win_Ratio) %>% 
  ungroup() 


df_t <- df %>% spread(key = Season, value = Win_Ratio) 


df_t$Team <- as.character(df_t$Team)



# inpute missing values with the mean
df <- df_t  %>% mutate_all(funs(ifelse(is.na(.), 0, .))) 

rownames(df) <- df$Team
df$Team <- NULL

```

## Hierarchical K-Means Clustering

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(proxy)
res.hk <- hkmeans(df, 3)
# Elements returned by hkmeans()


# Visualize the tree
fviz_dend(
  res.hk,
  cex = 0.4,
  palette = "Set1",
  rect = TRUE
)

# Visualize the hkmeans final clusters
fviz_cluster(res.hk,
             palette = "jco",
             repel = TRUE,
             ggtheme = theme_classic())


clusters <- as.data.frame(res.hk$cluster) %>% dplyr::rename(Cluster = "res.hk$cluster")
names <- as.data.frame(row.names(clusters)) %>% dplyr::rename(Team = "row.names(clusters)")

clusters$Team <- names$Team




```


# DTW clustering

Obtained clusters and their respective prototypes (centroids) shown as dashed lines.

```{r echo=FALSE, message=FALSE, warning=FALSE}

require("cluster")
hc_dtw <- tsclust(
  df,
  type = "h",
  k = 3L,
  distance = "dtw_basic",
  control = hierarchical_control(method = diana),
  args = tsclust_args(dist = list(window.size = 18L))
)

# By default, the dendrogram is plotted in hierarchical clustering
# plot(hc_dtw, xlab = "Teams", main = "Cluster Dendrogram of Stations", cex = 0.5)

# Obtained clusters and their respective prototypes (centroids) shown as dashed lines.
plot(hc_dtw, type = "sc")

# Prototype obtained by applying shape extraction to cluster 3
plot(hc_dtw, type = "centroids", clus = 1L)
plot(hc_dtw, type = "centroids", clus = 2L)
plot(hc_dtw, type = "centroids", clus = 3L)
```




# Analysis of CB for Teams' Clusters

## Analysis of Elite Teams


### NAMSI analysis of Cluster 2

```{r echo=FALSE, message=FALSE, warning=FALSE}
# auxiliar function that returns the denominator
my_fun <- function(n) {
  n <- n-1
  i <- 1
  result <- 0
  while (i < n-1) {
    result <- result + ((n - i) / (n - 1) - 0.5) ^ 2
    i <- i + 1
  }
  return(dplyr::data_frame(result = result))
}


# --------------

df <-
  data %>% group_by(Season, Team) %>% 
  summarise(
    Ws = sum(Game_Result == "G"),
    Ls = sum(Game_Result == "P"),
    Ts = sum(Game_Result == "E")
    ) %>% 
  mutate(Win_Ratio = Ws / (Ws + Ls + Ts))

tmp <- full_join(df, n_teams, by = "Season")


# filter data for elite teams
elite <- full_join(tmp, clusters, by = "Team") %>% filter(Cluster == 2)

# --------------


# number of teams
n_teams <- elite %>%
  group_by(Season, Team) %>%
  summarise(Total_Win_Ratio = sum(Win_Ratio)) %>%
  ungroup() %>%
  group_by(Season) %>%
  summarise(Count = n()) 

# denominator
denominator <- n_teams %>% 
  ungroup() %>%
  group_by(Season) %>%
  do(my_fun(n=.$Count)) %>% 
  mutate(denominator = sqrt(result))

# numerator
numerator <- elite %>%
  ungroup() %>%
  group_by(Season) %>%
  summarise(numerator = sqrt(sum((Win_Ratio - 0.5) ^ 2)))


NAMSI <- full_join(denominator, numerator, by = "Season") %>% 
  mutate(NAMSI = numerator/denominator)


NAMSI %>% ggplot(aes(Season, NAMSI)) +
  geom_line() +
  geom_smooth(span = 0.2)

```

### HICB analysis of Cluster 2

```{r echo=FALSE, message=FALSE, warning=FALSE}
# S squared
df <-
  data %>% group_by(Season, Team) %>%
  summarise(Total_Points = sum(Points)) %>%
  mutate(S_quared = Total_Points ^ 2)

# filter data for elite teams
elite <- full_join(df, clusters, by = "Team") %>% filter(Cluster == 2)


# number of teams
n_teams <- elite %>%
  group_by(Season, Team) %>%
  summarise(Total_Points = sum(Total_Points)) %>%
  ungroup() %>%
  group_by(Season) %>%
  summarise(Count = n())

# total points
total_points <- df %>%
  group_by(Season, Team) %>%
  summarise(Total_Points = sum(Total_Points)) %>%
  ungroup() %>%
  group_by(Season) %>%
  summarise(Total_Points_Season = sum(Total_Points))

tmp <- full_join(elite, n_teams, by = "Season")
tmp <- full_join(tmp, total_points, by = "Season")

HICB <- tmp %>%
  ungroup() %>%
  group_by(Season) %>%
  summarise(HICB = sum(Count / Count) * 100 * (sum((
    Total_Points / Total_Points_Season
  ) ^ 2)))

HICB %>% ggplot(aes(Season, HICB)) +
  geom_line()+
  geom_smooth(span = 0.2)

```

## Analysis of Regular Teams

### NAMSI analysis of Cluster 1

```{r echo=FALSE, message=FALSE, warning=FALSE}
# auxiliar function that returns the denominator
my_fun <- function(n) {
  n <- n-1
  i <- 1
  result <- 0
  while (i < n-1) {
    result <- result + ((n - i) / (n - 1) - 0.5) ^ 2
    i <- i +1
  }
  return(dplyr::data_frame(result = result))
}


# --------------

df <-
  data %>% group_by(Season, Team) %>% 
  summarise(
    Ws = sum(Game_Result == "G"),
    Ls = sum(Game_Result == "P"),
    Ts = sum(Game_Result == "E")
    ) %>% 
  mutate(Win_Ratio = Ws / (Ws + Ls + Ts))

tmp <- full_join(df, n_teams, by = "Season")


# filter data for elite teams
elite <- full_join(tmp, clusters, by = "Team") %>% filter(Cluster == 1)

# --------------


# number of teams
n_teams <- elite %>%
  group_by(Season, Team) %>%
  summarise(Total_Win_Ratio = sum(Win_Ratio)) %>%
  ungroup() %>%
  group_by(Season) %>%
  summarise(Count = n()) 

# denominator
denominator <- n_teams %>% 
  ungroup() %>%
  group_by(Season) %>%
  do(my_fun(n=.$Count)) %>% 
  mutate(denominator = sqrt(result))

# numerator
numerator <- elite %>%
  ungroup() %>%
  group_by(Season) %>%
  summarise(numerator = sqrt(sum((Win_Ratio - 0.5) ^ 2)))

NAMSI <- full_join(denominator, numerator, by = "Season") %>% 
  mutate(NAMSI = ifelse(denominator == 0, 0, numerator/denominator))

mean <- mean(NAMSI$NAMSI)

NAMSI <- NAMSI %>% 
  mutate(NAMSI = ifelse(denominator == 0, mean, numerator/denominator))


NAMSI %>% ggplot(aes(Season, NAMSI)) +
  geom_line() +
  geom_smooth(span = 0.2)

```

### HICB analysis of Cluster 1

```{r echo=FALSE, message=FALSE, warning=FALSE}
# S squared
df <-
  data %>% group_by(Season, Team) %>%
  summarise(Total_Points = sum(Points)) %>%
  mutate(S_quared = Total_Points ^ 2)

# filter data for elite teams
elite <- full_join(df, clusters, by = "Team") %>% filter(Cluster == 1)


# number of teams
n_teams <- elite %>%
  group_by(Season, Team) %>%
  summarise(Total_Points = sum(Total_Points)) %>%
  ungroup() %>%
  group_by(Season) %>%
  summarise(Count = n())

# total points
total_points <- df %>%
  group_by(Season, Team) %>%
  summarise(Total_Points = sum(Total_Points)) %>%
  ungroup() %>%
  group_by(Season) %>%
  summarise(Total_Points_Season = sum(Total_Points))

tmp <- full_join(elite, n_teams, by = "Season")
tmp <- full_join(tmp, total_points, by = "Season")

HICB <- tmp %>%
  ungroup() %>%
  group_by(Season) %>%
  summarise(HICB = sum(Count / Count) * 100 * (sum((
    Total_Points / Total_Points_Season
  ) ^ 2)))

HICB %>% ggplot(aes(Season, HICB)) +
  geom_line()+
  geom_smooth(span = 0.2)

```



## Analysis of Poor Teams

### NAMSI analysis of Cluster 3

```{r echo=FALSE, message=FALSE, warning=FALSE}
# auxiliar function that returns the denominator
my_fun <- function(n) {
  n <- n-1
  i <- 1
  result <- 0
  while (i < n-1) {
    result <- result + ((n - i) / (n - 1) - 0.5) ^ 2
    i <- i +1
  }
  return(dplyr::data_frame(result = result))
}


# --------------

df <-
  data %>% group_by(Season, Team) %>% 
  summarise(
    Ws = sum(Game_Result == "G"),
    Ls = sum(Game_Result == "P"),
    Ts = sum(Game_Result == "E")
    ) %>% 
  mutate(Win_Ratio = Ws / (Ws + Ls + Ts))

tmp <- full_join(df, n_teams, by = "Season")


# filter data for elite teams
elite <- full_join(tmp, clusters, by = "Team") %>% filter(Cluster == 3)

# --------------


# number of teams
n_teams <- elite %>%
  group_by(Season, Team) %>%
  summarise(Total_Win_Ratio = sum(Win_Ratio)) %>%
  ungroup() %>%
  group_by(Season) %>%
  summarise(Count = n())  

# denominator
denominator <- n_teams %>% 
  ungroup() %>%
  group_by(Season) %>%
  do(my_fun(n=.$Count)) %>% 
  mutate(denominator = sqrt(result))

# numerator
numerator <- elite %>%
  ungroup() %>%
  group_by(Season) %>%
  summarise(numerator = sqrt(sum((Win_Ratio - 0.5) ^ 2)))


NAMSI <- full_join(denominator, numerator, by = "Season") %>% 
  mutate(NAMSI = ifelse(denominator == 0, 0, numerator/denominator))


NAMSI %>% ggplot(aes(Season, NAMSI)) +
  geom_line() +
  geom_smooth(span = 0.2)

```


### HICB analysis of Cluster 3

```{r echo=FALSE, message=FALSE, warning=FALSE}
# S squared
df <-
  data %>% group_by(Season, Team) %>%
  summarise(Total_Points = sum(Points)) %>%
  mutate(S_quared = Total_Points ^ 2)

# filter data for elite teams
elite <- full_join(df, clusters, by = "Team") %>% filter(Cluster == 3)


# number of teams
n_teams <- elite %>%
  group_by(Season, Team) %>%
  summarise(Total_Points = sum(Total_Points)) %>%
  ungroup() %>%
  group_by(Season) %>%
  summarise(Count = n())

# total points
total_points <- df %>%
  group_by(Season, Team) %>%
  summarise(Total_Points = sum(Total_Points)) %>%
  ungroup() %>%
  group_by(Season) %>%
  summarise(Total_Points_Season = sum(Total_Points))

tmp <- full_join(elite, n_teams, by = "Season")
tmp <- full_join(tmp, total_points, by = "Season")

HICB <- tmp %>%
  ungroup() %>%
  group_by(Season) %>%
  summarise(HICB = sum(Count / Count) * 100 * (sum((
    Total_Points / Total_Points_Season
  ) ^ 2)))

HICB %>% ggplot(aes(Season, HICB)) +
  geom_line()+
  geom_smooth(span = 0.2)

```


# Kendall

## Based on Total points

```{r echo=FALSE, message=FALSE, warning=FALSE}

#-------- EXAMPLE -------#
tmp1 <- data_frame(Var1 = c("0.0", "0.5", "0.75"))
tmp2 <- data_frame(Var2 = c("0.1", "0.0", "0.2"))

tmp1$Var1 <- as.numeric(tmp1$Var1)
tmp2$Var2 <- as.numeric(tmp2$Var2)

# a <- as.numeric(factor(tmp1$Var1, levels = c("A", "B", "C", "D", "E", "F")))
# <- as.numeric(factor(tmp2$Var2, levels = c("A", "B", "C", "D", "E", "F")))
m <- cbind(tmp1, tmp2)
cor(m, method = "kendall", use = "pairwise")

#-------- END OF EXAMPLE -------#

```

## Based on SD

```{r echo=FALSE, message=FALSE, warning=FALSE}
df <-
  data %>% group_by(Season, Team) %>% 
  summarise(
    Ws = sum(Game_Result == "G"),
    Ls = sum(Game_Result == "P"),
    Ts = sum(Game_Result == "E")
    ) %>% 
  mutate(Win_Ratio = Ws / (Ws + Ls + Ts))

# number of teams
n_teams <- data %>%
  group_by(Season, Team) %>%
  summarise(Total_Points = sum(Points)) %>%
  ungroup() %>%
  group_by(Season) %>%
  summarise(Count = n())

tmp <- full_join(df, n_teams, by = "Season")

tmp <- tmp %>%
  ungroup() %>%
  group_by(Season, Team) %>%
  mutate(SD = (Win_Ratio - 0.5) ^ 2) %>%
  ungroup() %>%
  group_by(Season, Team) %>%
  summarise(SD = sqrt(mean(SD)), Teams_Count = sum(Count/Count)) %>% 
  dplyr::select(c(-Teams_Count)) %>% 
  filter(Season >= 1941) 

tmp <- tmp %>%
  spread(key = Season, value = SD)

# cluster1
cluster1 <- full_join(tmp, clusters, by = "Team") %>%
  filter(Cluster == 1) %>%
  select(c(-Cluster))

# cluster2
cluster2 <- full_join(tmp, clusters, by = "Team") %>%
  filter(Cluster == 2) %>%
  select(c(-Cluster))

# cluster3
cluster3 <- full_join(tmp, clusters, by = "Team") %>%
  filter(Cluster == 3) %>%
  select(c(-Cluster))

# visualize the clusters
DT::datatable(cluster1, style = "bootstrap", fillContainer = TRUE)
DT::datatable(cluster2, style = "bootstrap", fillContainer = TRUE)
DT::datatable(cluster3, style = "bootstrap", fillContainer = TRUE)


cluster1 <- cluster1 %>% mutate_all(funs(ifelse(is.na(.), 0, .)))
cluster2 <- cluster2 %>% mutate_all(funs(ifelse(is.na(.), 0, .)))
cluster3 <- cluster3 %>% mutate_all(funs(ifelse(is.na(.), 0, .)))

kendall_fun <- function(df) {
  i <- 2
  j <- 3
  result <- c()
  while (j <= ncol(df)) {
    m <- cbind(df[, i], df[, j])
    kendall <- cor(m, method = "kendall", use = "pairwise")[1, 2]
    result <- c(result, kendall)
    j <- j + 1
    i <- i + 1
  }
  return(result)
}

result <- kendall_fun(cluster1)
season <- c(1942:2017)
kendall_all <- data_frame(Kendall = result, Season = season)
kendall_all %>% ggplot(aes(x = Season, y = Kendall)) +
  geom_smooth(span = 0.2, size = 2) +
  geom_line(size = 1) +
  labs(title = "Cluster 1")

result <- kendall_fun(cluster2)
season <- c(1942:2017)
kendall_all <- data_frame(Kendall = result, Season = season)
kendall_all %>% ggplot(aes(x = Season, y = Kendall)) +
  geom_smooth(span = 0.2, size = 2) +
  geom_line(size = 1) +
  labs(title = "Cluster 2")

result <- kendall_fun(cluster3)
season <- c(1942:2017)
kendall_all <- data_frame(Kendall = result, Season = season)
kendall_all %>% ggplot(aes(x = Season, y = Kendall)) +
  geom_smooth(span = 0.2, size = 2) +
  geom_line(size = 1) +
  labs(title = "Cluster 3")


```


# Causal Impact Analysis

```{r echo=FALSE, message=FALSE, warning=FALSE}

df <-
  data %>% group_by(Season, Team) %>%
  summarise(
    Ws = sum(Game_Result == "G"),
    Ls = sum(Game_Result == "P"),
    Ts = sum(Game_Result == "E")
  ) %>%
  filter(Season >= 1942) %>%
  summarise(Mean_Ts = mean(Ts)) %>% 
  mutate(Mean_Ts_Norm = Mean_Ts/max(Mean_Ts))


# cluster1
cluster1 <- full_join(tmp, clusters, by = "Team") %>%
  filter(Cluster == 1) %>%
  select(c(-Cluster))

cluster1 <- cluster1 %>% mutate_all(funs(ifelse(is.na(.), 0, .)))


y <- kendall_fun(cluster1)
x1 <- as.numeric(df$Mean_Ts_Norm)
time.points <- seq.Date(as.Date("1941-01-01"), by = "year", length.out = 77)

data <- zoo(cbind(y, x1), time.points)

pre.period <- as.Date(c("1941-01-01", "1996-03-11"))
post.period <- as.Date(c("1997-01-01", "2017-03-11"))

impact <- CausalImpact(data, pre.period, post.period)
plot(impact)



# cluster2
cluster2 <- full_join(tmp, clusters, by = "Team") %>%
  filter(Cluster == 2) %>%
  select(c(-Cluster))

cluster2 <- cluster2 %>% mutate_all(funs(ifelse(is.na(.), 0, .)))


y <- kendall_fun(cluster2)
x1 <- as.numeric(df$Mean_Ts_Norm)
time.points <- seq.Date(as.Date("1941-01-01"), by = "year", length.out = 77)

data <- zoo(cbind(y, x1), time.points)

pre.period <- as.Date(c("1941-01-01", "1996-03-11"))
post.period <- as.Date(c("1997-01-01", "2017-03-11"))

impact <- CausalImpact(data, pre.period, post.period)
plot(impact)




# cluster3
cluster3 <- full_join(tmp, clusters, by = "Team") %>%
  filter(Cluster == 3) %>%
  select(c(-Cluster))

cluster3 <- cluster3 %>% mutate_all(funs(ifelse(is.na(.), 0, .)))


y <- kendall_fun(cluster3)
x1 <- as.numeric(df$Mean_Ts_Norm)
time.points <- seq.Date(as.Date("1941-01-01"), by = "year", length.out = 77)

data <- zoo(cbind(y, x1), time.points)

pre.period <- as.Date(c("1941-01-01", "1996-03-11"))
post.period <- as.Date(c("1997-01-01", "2017-03-11"))

impact <- CausalImpact(data, pre.period, post.period)
plot(impact)
```





